Towards a Complete Model of Mutation Selection Balance in Humans

### 2023-05-06 manuscript/manuscript.tex
ms updates

\section*{Results}
\subsection*{Theory}

Similar to previous work, our method characterizes the negative selection
process through its indirect effects on neighboring levels of pairwise
diversity. While pairwise diversity is a less-informative statistic than the
full site frequency spectrum, it is far easier to derive models of how
selection alters expected pairwise coalescence times at linked sites than to
describe the entire genealogy. Statistical methods then fit observed diversity
to the expectation of a particular theoretic model of the linked selection
process. Consequently, statistical inferences about selection may be incorrect
when the underlying theory inaccurately predicts average pairwise coalescence
times.

Our work is motivated by a well-known limitation of classic background
selection theory: it is only accurate for strongly deleterious mutations. Given
that statistical methods rely on background selection theory to estimate the
mutation rate and distribution of fitness effects of deleterious mutations, it
is important to characterize and remedy these shortcomings of the model.
However, this is

It is well-known that the approximations that underpin classic background
selection theory
\parencite{Charlesworth1993-gb,Nordborg1996-nq,Hudson1994-oh,Hudson1995-pt}
break down when selection against deleterious mutations is weak, and the
reduction in diversity predicted by this model is incorrect
\parencite{McVean2000-bt,Good2013-lp,Gordo2002-dr}. Before introducing how th

>>>>

\section*{Results}

Similar to previous work, our method characterizes the negative selection
process through the signal left on pairwise diversity at linked sites. Although
pairwise diversity is a less-informative statistic than the full site frequency
spectrum, it is much easier to derive models of how selection alters expected
pairwise coalescence times at linked sites than to describe the entire
genealogy. Spatial patterns of pairwise diversity are highly variable across
eukaryotic genomes, as linked selection leads diversity to become correlated
with spatial heterogeneity in recombination rates and the density of conserved
sites. Statistical methods then fit observed diversity to the expected level
under a particular theoretic model of linked selection processes. Consequently,
statistical inferences about selection may be incorrect when the underlying
theory inaccurately predicts average pairwise coalescence times.

Our work is motivated by a well-known limitation of classic background
selection theory: it is only accurate for strongly deleterious mutations
\parencite{Charlesworth1993-gb,McVean2000-bt,Good2013-lp,Gordo2002-dr}. Given
statistical methods that fit genome-wide diversity to linked selection
processes rely on the classic BGS model to estimate the distribution of fitness
effects and rate of deleterious mutations, it is important to characterize and
remedy these shortcomings. However, extending background selection theory to
model weak negative selection is challenging, and it is worthwhile to develop
some intuition for how the classic background selection works and why this is
the case.

The classic background selection model \parencite{Charlesworth1993-gb}

First, the frequency trajectory of a weakly selected allele is
predominantly determined by stochastic evolutionary processes. When the changes
due to stochasticity are large, deterministic approximations that work well
under strong selection no longer capture selection dynamics. Second, weakly
selected alleles are impacted by two sources of evolutionary stochasticity:
genetic drift (i.e. non-heritable differences in offspring number and Mendelian
segregation) and genetic draft, the stochastic changes in allele frequency
caused by selection on the same genetic background
\parencite{Neher2013-dz,McVean2000-bt}.

Moreover, when the magnitude of stochastic changes can exceed the deterministic
changes caused by selection, mutations that decrease fitness now have a chance
of fixation \parencite{Kimura1962-su,Malecot1948-zv}. Because there is a steady
influx of weakly deleterious mutations, each with a non-zero chance of
substitution, there must be a corresponding substitution process. The rate of
this deleterious substitution process (e.g. ``Muller's ratchet") is

By contrast, when selection is assumed to be strong,
deterministic models of mutation and selection that ignore these sources of
stochasticity adequately capture the dynamics. Since weakly deleterious
mutations have some chance of fixation,

\subsection*{Theory}
